### 2014-12-03 docs/ms.Rmd
Re-draft intro, begin restructuring to focus more broadly on reproducibility

# The transparency of methods-reporting in studies of biological colouration

@@

## Abstract
Over the last 50 years or so, our enhanced capacity for quantitative colour measurement has accelerated understanding of the evolution of biological colouration. However, while spectrometry and digital photography are inexpensive, accessible and efficient, details of exactly how data are collected and analysed are not reported consistently in the literature. Similarly, estimates of animals’ visual capacity are widely applied to colour data, but details on assumptions and settings in analyses are not reported completely. To quantify the variability of methods reporting in studies of colouration, we reviewed papers published in 2013 in 22 leading journals. Using the search terms colour* OR color* OR spectra* in the title or abstract, we identified 60 papers in which colour had been measured and analysed. We extracted the methods and noted the details included. We found a great deal of variability in the level of detail in which colour methods is reported. In response we have outlined explicit recommendations on the important methodological details required to ensure studies are transparent and repeatable. We hope that this will help to standardize methods reporting in the field of biological colouration paving the way for direct comparison across studies.

## Introduction
Generations of biologists have endeavored to explain the mechanisms and functions of animal and plant colouration [@thayer_concealing_1909; @poulton_colours_1890; @wallace_natural_1891]. For these biologists, identifying best practices in measuring colour has been a great challenge. Over the past 50 years, biologists have developed and adapted techniques for the objective measurement of colouration that are now widespread. These include spectrometry, photography, and visual modelling.

The application of spectrometry to questions of biological colouration revolutionized the field [@dyck_determination_1966]. Before its widespread use, researchers assigned categorical colour descriptions, arbitrary rankings or scores [e.g. @aldrich_ecogeographic_1991; @bailey_latitudinal_1978; @wheelwright_colors_1985] or matched paint swatches or munsell chip ratings based on their own visual assessment [e.g. @burtt_analysis_1986; @key_temperature_1954]. Although no doubt applied carefully, these techniques limited progress. For example, many researchers designed colour assessment methods specifically for their projects, rendering data incomparable across studies [e.g. @aldrich_ecogeographic_1991; @bailey_latitudinal_1978]. This is especially problematic when different measurement techniques are used on the same species [e.g. house finch _Carpodacus mexicanus_; @aldrich_origin_1978; @hill_female_1990]. Likewise, because human vision is inherently different to most other animals [@bennett_sexual_1994; @douglas_spectral_2014], its use has lead to erroneous conclusions [@bennett_sexual_1994; @endler_measurement_1990].

@@

## Ideal reporting in colour studies

We developed a list of information we deemed critical to be reported for reproducibility of colour methods (Table 1) and analysis (Table 2), loosely based on Andersson and Prager [-@andersson_quantifying_2006] and Montgomerie [@montgomerie_analyzing_2006]. These included quantitative methods including spectrometric and photographic techniques [as in @kemp_ornamental_2008; @chiao_visualization_2009] but not qualitative methods such as the use of Munsell chips or paint swatches [as in @hogstad_gloger_2009]. For visual modeling techniques we characterized a set of criteria required for reproducing the analysis (Table 2). Much of these criteria specifically targeted the necessary parameters required to be explicitly defined by the user during analysis. These are essential if such analyses are to be reproduced. These criteria provided that baseline against which to hold our results.

## What is reporting like in the literature?

To determine the current state of methods reporting in the field we searched for papers from 2013 in 22 leading journals (_American Journal of Botany_, _The American Naturalist_, _Animal Behaviour_, _Behavioral Ecology_, _Behavioral Ecology and Sociobiology_, _Biological Journal of the Linnean Society_, _Biology Letters_, _Current Zoology_, _Ecology_, _Ecology and Evolution_, _Ecology Letters_, _Ethology_, _Evolution_, _Functional Ecology_, _Journal of Ecology_, _The Journal of Evolutionary Biology_, _The Journal of Experimental Biology_, _Naturwissenschaften_, _New Phytologist_, _Oikos_, _PLoS ONE_, _Proceedings of the Royal Society B: Biological Sciences_). On each of the journals' homepages, we used the Boolean terms 'colour*', 'color*' or 'spectra*' to search the title or abstract. We read the titles and abstracts of each paper and included only those papers that used either a spectrometer or camera to quantify colouration. We excluded papers review papers, methodological papers, papers quantifying colour patterns rather than the chromatic properties of a colour patch, and papers taking microspectrophotometric measurments of retinal absorbance. To review the large number of papers returned from our search, journals were haphazardly divided up between the authors. In order to reduce the risk of observer bias in our assesment, each paper was read and reassessed by three authors. Any discrepancies between assessor scores were discussed and resolved prior to analysis. We then tabulated what details of the methods were reported against our criteria. The data along with analysis script can be found at: https://github.com/daniel1noble/colsci_rep (TODO: GIVE IT A DOI?). We have kept the papers used in our dataset annoymous as our aim is to make a general statement oabout methods reporting in the field.

Our search returned a total of 216 papers, however, only `r nrow(coldat)` met our criteria and were included in the final analysis. Overall, `r specs+both` studies used a spectrometer to measure colour, while `r cam+both` used a camera. Of these studies, `r both` quantified colour using both a spectrometer and a camera. Reporting of the make and model of the spectrometer and camera was high (Figure 1, spectrometers = `r round(specModel, digits = 2)`% and cameras = `r round(camModel, digits = 2)`%), however, only `r round(propCam_fin["prop_light_source"], digits = 2)`%  and `r round(propSpec_fin["prop_light_source"], digits = 2)`% of camera and spectrometer studies reported a light source. In addition, the number of reflectance spectra and pixels averaged were only reported in `r round(propSpec_fin["prop_specpix_avg"], digits = 2)`% and `r round(propCam_fin["prop_specpix_avg"], digits = 2)`% of studies, respectively. Surprisingly, dark standards were only reported in `r round(propSpec_fin["prop_dark_std"], digits = 2)`% of the spectrometer studies, while white standards were reported in `r round(propSpec_fin["prop_white_stdspec"], digits = 2)`%. Studies quantifying spectral reflectance often did not report the integration time of the spectrophotmeter (`r round(propSpec_fin["prop_int_time"], digits = 2)`%), or the angle of the optical probe relative to the surface (`r round(propSpec_fin["prop_spec_angle"], digits = 2)`%) or distance (`r round(propSpec_fin["prop_spec_dist"], digits = 2)`%) the probe was held from the measurement surface.

@@

##Conclusions

@@

While all criteria are important, some deserve special attention. In particular, we found that the types of light sources, numbers of pixels or spectra averaged, and types of standards used (i.e. white and black reference samples) are under-reported, each of which is critical to repeatability and cross-comparison. Similarly, the distance and angle of the optical probe relative to the measurment surface and collector should be standard reporting for all studies using spectrophometry, as variation in these factors can lead to dramatically different results in some cases (REFS). Visual models were regularly used to analyse spectral data, however, there were many important aspects of the analysis that were seldom reported. These included the type of quantum catch used, whether photoreceptor adaptation was modelled and the type of noise used in receptor-noise models. This is problematic because... Also, photoreceptor densities should be reported where possible even if using standard avian photorecptor densities because.... These all have important consequences on downstream analysis and the reproducibility of colour signal quantification [e.g. @kemp_ornamental_2008; @santos_comparison_2007]. We also recommend the presentation of spectral reflectance curves where possible. These figures allow for the rapid assesment of the nature and quality of the data, independent of downstream processing (if the conditions under which they were captured is comprehensively reported. We found that only `r round(propSpec["refl_fig"], digits = 2)`% of studies using spectral reflectance data had a figure showing these curves. When possible, we also recommend that authors justify data capture and analytical approaches - something few studies we investigated explicitly outlined.

>>>>

# Reproducible research in biological colour science

@@

## Introduction

Studies of biological colouration have produced insights into fundamental evolutionary and ecological processes, and research into colour traits is a rapidly growing, interdisciplinary field []. The resurgent interest in biological colouration has in large part been driven by the increased availability of key technologies, including spectrometry and photography, and concurrent developments in methods for analysing colour data, such as visual models []. While such advances are extremely positive for the field, the rapid rate of progression means that a deep understanding of common methods is increasingly beyond the grasp of the average researcher. Indeed, the subtle complexity of these techniques [e.g. the 3-7 parameters that define a typical visual model; ] is often overlooked by empiricists, which leads to critical methodological information often not being reported (Fig. 1); thus limiting a study's utility.

Replication and transparency lie at the heart of the scientific enterprise. True replication of individual studies based on independantly collected data is the gold standard in cumulative science. This may often be unfeasable however, espescially in field-based research, which includes a great deal of research on colour-traits []. In contrast, reproducability - the re-analysis of a study's data by independant researchers - has been presented as a readily attainable minimum standard [@peng_reproducible_2006; @peng_reproducible_2011]. A study may be considered truly reproducible when it satifies three broad criteria: (i) methods are reported completely, (ii) data are publicly archived and available, and (iii) analysis code is made publicly available, such that the chain of modification of raw data is documented and preserved. While completely reproducible research [e.g. @fitzjohn_much_2014] is a laudable goal, the considerable demands on researchers means that it will often, in practice, be unattanable. Nevertheless, even partial reproducibility is of tremendous value as it allows greater comparability between studies, and may serve as a launching point for future research.

Our aim here to explore current levels of reproducibility in studies of biological colouration, and to outline ways in which they may be improved in line with points (i) - (iii) above. We begin by briefly outlining some of the most common methods for studying biological colouration. We then present a set of guidelines for the reporting of methods for the capture and analysis of colour data, and we assess a subset of the literature against these criteria to assess the current state of methods reporting in colour science. We also quantify the availability of publicly archived data and code, and suggest some tools...

## Measuring and analysing colour

todo: hack this back

Generations of biologists have endeavored to explain the mechanisms and functions of animal and plant colouration [@thayer_concealing_1909; @poulton_colours_1890; @wallace_natural_1891], and identifying best practices in measuring colour has been a great challenge. The application of spectrometry to questions of biological colouration revolutionized the field [@dyck_determination_1966]. Before its widespread use, researchers assigned categorical colour descriptions, arbitrary rankings or scores [e.g. @aldrich_ecogeographic_1991; @bailey_latitudinal_1978; @wheelwright_colors_1985] or matched paint swatches or munsell chip ratings based on their own visual assessment [e.g. @burtt_analysis_1986; @key_temperature_1954]. Although no doubt applied carefully, these techniques limited progress. For example, many researchers designed colour assessment methods specifically for their projects, rendering data incomparable across studies [e.g. @aldrich_ecogeographic_1991; @bailey_latitudinal_1978]. This is especially problematic when different measurement techniques are used on the same species [e.g. house finch _Carpodacus mexicanus_; @aldrich_origin_1978; @hill_female_1990]. Likewise, because human vision is inherently different to most other animals [@bennett_sexual_1994; @douglas_spectral_2014], its use has lead to erroneous conclusions [@bennett_sexual_1994; @endler_measurement_1990].

@@

## Assessing reproducibility

To determine the current state of reproducibility in the field we searched for papers from 2013 in 22 leading journals (_American Journal of Botany_, _The American Naturalist_, _Animal Behaviour_, _Behavioral Ecology_, _Behavioral Ecology and Sociobiology_, _Biological Journal of the Linnean Society_, _Biology Letters_, _Current Zoology_, _Ecology_, _Ecology and Evolution_, _Ecology Letters_, _Ethology_, _Evolution_, _Functional Ecology_, _Journal of Ecology_, _The Journal of Evolutionary Biology_, _The Journal of Experimental Biology_, _Naturwissenschaften_, _New Phytologist_, _Oikos_, _PLoS ONE_, _Proceedings of the Royal Society B: Biological Sciences_). On each of the journals' homepages, we used the Boolean terms 'colour*', 'color*' or 'spectra*' to search the title or abstract. We read the titles and abstracts of each paper and included only those papers that used either a spectrometer or camera to quantify colouration. We excluded papers review papers, methodological papers, papers quantifying colour patterns rather than the chromatic properties of a colour patch, and papers taking microspectrophotometric measurments of retinal absorbance. To review the large number of papers returned from our search, journals were haphazardly divided up between the authors. In order to reduce the risk of observer bias in our assesment, each paper was read and reassessed by three authors. Any discrepancies between assessor scores were discussed and resolved prior to analysis. We then tabulated what details of the methods were reported against our criteria. The data along with analysis script can be found at: https://github.com/daniel1noble/colsci_rep (TODO: GIVE IT A DOI?). We have kept the papers used in our dataset annoymous as our aim is to make a general statement about methods reporting in the field.

## Methods reporting in colour studies

_Guidelines for methods reporting_

todo: expand table discussion

We developed a list of information we deemed critical to be reported for reproducibility of colour methods (Table 1) and analysis (Table 2), loosely based on Andersson and Prager [-@andersson_quantifying_2006] and Montgomerie [@montgomerie_analyzing_2006]. These included quantitative methods including spectrometric and photographic techniques [as in @kemp_ornamental_2008; @chiao_visualization_2009] but not qualitative methods such as the use of Munsell chips or paint swatches [as in @hogstad_gloger_2009]. For analytical techniques we characterized a set of criteria required for reproducing the analysis (Table 2). Much of these criteria specifically targeted the necessary parameters required to be explicitly defined by the user during analysis. These are essential if such analyses are to be reproduced. These criteria provided that baseline against which to hold our results. Our search returned a total of 216 papers, however, only `r nrow(coldat)` met our criteria and were included in the final analysis.

_The curent state of methods reporting in the literature_

Overall, `r specs+both` studies used a spectrometer to measure colour, while `r cam+both` used a camera. Of these studies, `r both` quantified colour using both a spectrometer and a camera. Reporting of the make and model of the spectrometer and camera was high (Figure 1, spectrometers = `r round(specModel, digits = 2)`% and cameras = `r round(camModel, digits = 2)`%), however, only `r round(propCam_fin["prop_light_source"], digits = 2)`%  and `r round(propSpec_fin["prop_light_source"], digits = 2)`% of camera and spectrometer studies reported a light source. In addition, the number of reflectance spectra and pixels averaged were only reported in `r round(propSpec_fin["prop_specpix_avg"], digits = 2)`% and `r round(propCam_fin["prop_specpix_avg"], digits = 2)`% of studies, respectively. Surprisingly, dark standards were only reported in `r round(propSpec_fin["prop_dark_std"], digits = 2)`% of the spectrometer studies, while white standards were reported in `r round(propSpec_fin["prop_white_stdspec"], digits = 2)`%. Studies quantifying spectral reflectance often did not report the integration time of the spectrophotmeter (`r round(propSpec_fin["prop_int_time"], digits = 2)`%), or the angle of the optical probe relative to the surface (`r round(propSpec_fin["prop_spec_angle"], digits = 2)`%) or distance (`r round(propSpec_fin["prop_spec_dist"], digits = 2)`%) the probe was held from the measurement surface.

@@

## The public availability of data and code

todo: results here

_Tools for reproducible research_

Data - archiving (figshare, dryad), raw data preferable to processed data, concerns over data sharing, embargos are fine just get it archived.
Code - benefits of code over point-and-click, open source shit (R, pavo, things), github, sourcefourge etc.

## Conclusions

@@

While all criteria are important, some deserve special attention. In particular, we found that the types of light sources, numbers of pixels or spectra averaged, and types of standards used (i.e. white and black reference samples) are under-reported, each of which is critical to repeatability and cross-comparison. Similarly, the distance and angle of the optical probe relative to the measurment surface and collector should be standard reporting for all studies using spectrophometry, as variation in these factors can lead to dramatically different results in some cases (REFS). Visual models were regularly used to analyse spectral data, however, there were many important aspects of the analysis that were seldom reported. These included the type of quantum catch used, whether photoreceptor adaptation was modelled and the type of noise used in receptor-noise models. This is problematic because... Also, photoreceptor densities should be reported where possible even if using standard avian photorecptor densities because.... These all have important consequences on downstream analysis and the reproducibility of colour signal quantification [e.g. @kemp_ornamental_2008; @santos_comparison_2007]. We also recommend the presentation of spectral reflectance curves where possible. These figures allow for the rapid assesment of the nature and quality of the data, independent of downstream processing (if the conditions under which they were captured is comprehensively reported. We found that only `r round(propSpec_fin["prop_refl_fig"], digits = 2)`% of studies using spectral reflectance data had a figure showing these curves. When possible, we also recommend that authors justify data capture and analytical approaches - something few studies we investigated explicitly outlined.
